{\bf Practical Concurrency Testing}

### 2017-04-25 proposal/talk/proposal.tex
finish slides??
finalize slides i think

\begin{document}
%\renewcommand{\inserttotalframenumber}{39}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	Possible interleavings represented as a tree
	\begin{itemize}
		\item Node: Intermediate execution state
		\item Edge: State transition from executing a thread
	\end{itemize}
	%\linegap

	%State space size depends on which {\bf preemption points} are used.
\end{frame}

@@

	Kernel (``P3'')
	\begin{itemize}
		%\item UNIX-like educational kernel specification
		\item Students implement scheduler, synchronization, VM %from scratch %% trying not to be too arrogant ._.
		\item UNIX-like process lifecycle ({\tt fork}, {\tt exec}, {\tt vanish}, {\tt wait})
		%\item Teams of 2
		\item Testing with MC requires manual annotation {\em [Blum '12]}
	\end{itemize}
	\pause
	\linegap

@@

	Results (as of spring 2016)
	\begin{itemize}
		\item 47/90 groups volunteered
		\item 85 concurrency bugs found; 53 which students then fixed
		\item No visible impact on ultimate grades...
	\end{itemize}
\end{frame}

@@

	Kernel (``threads'')
	\begin{itemize}
		\item Students implement priority scheduling, sleep
		\item Basecode provides round-robin scheduler, synchronization
	\end{itemize}
	\linegap

@@

\begin{frame}{Pintos \& Landslide}
	Existing support:
	\begin{itemize}
		\item Lockset/vector-clock analyses extended to interrupt-disabling
		\item Automatically inserting (most) kernel annotations
		%\item (various engineering challenges) % palloc heap, non-direct-mapped kernelspace, skipping driver init loops
		\item Often requires manual intervention
	\end{itemize}
	% "Works enough for OOPSLA but not enough for the studence"
	\linegap

@@

\begin{frame}{Deliverables}
	Pebbles
	\begin{itemize}
		\item Discuss factors keeping Landslide from improving grades
		%\item Discuss how to make P3 and Landslide compromise to work together
		\item Study accuracy of automation heuristics
	\end{itemize}
	\linegap

	Pintos
	\begin{itemize}
		\item Open-source port of Landslide to Bochs
		\item 2 semesters of user study in Pintos class(es)
		\item Compare incidence/nature of bugs found to Pebbles
	\end{itemize}
	\linegap
\end{frame}

@@

	MCing a transactional program
	\begin{itemize}
		\item Simulate TM, interleaving transactions, rolling-back upon conflicts
			\begin{itemize}
				\item slow; invites state space explosion
			\end{itemize}
		\item Enforce atomicity via scheduler, anticipate aborts via analysis
			\begin{itemize}
				\item (under HTM, all transactions might abort)
			\end{itemize}
		\item Suffices to enumerate all {\em observationally-equivalent} schedules.
	\end{itemize}
\end{frame}

\begin{frame}{Concurrency model}
	What are a transactional program's {\em observable} behaviours?
	\begin{itemize}
		\item Transactional writes, if successful, appear atomically %to other threads
		\begin{itemize}
			\item Equivalent to a lock protecting that state
		\end{itemize}
			\pause
		\item Simultaneous transactions succeed only if no conflicts
		\begin{itemize}
			\item Equivalent (modulo performance) to serialize independent transactions
		\end{itemize}
			\pause
		\item Failing transactions' effects rolled-back with no trace
		\begin{itemize}
			\item Treat each \xbegin as a failure injection point
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Failure Injection - 2 dimensions of concurrency}
	% TODO - pict of mario man
\end{frame}

@@

\begin{frame}{Timeline}
	\begin{itemize}
		\item 2017 Mar-Aug: Port Landslide to Bochs for Pintos
		\item 2017 Mar-Aug: Transactional memory
		\item 2017 Aug-Oct: Pintos user study
		\item 2017 Oct-Nov: Pebbles user study (again)
		\item 2017 Dec:     Start writing thesis
		\item 2018 Jan-Apr: More writing \& miscellany
		\item 2018 Feb-Mar: Pintos \& Pebbles (again)
		\item 2018 Apr-May: Defend \& graduate
	\end{itemize}
\end{frame}

@@

\begin{frame}{Interleaving t'xns non-atomically $\rightarrow$ state space explosion}
					% TODO: need a bonus slide to justify why this
					% imagine a pair of transactions each with multiple conflicting accesses
					% they should be treated as atomic
\end{frame}

\begin{frame}{STM: Judging when aborts are possible}
	% TODO
\end{frame}

@@

@@

>>>>

\begin{document}
\renewcommand{\inserttotalframenumber}{39}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	Possible interleavings represented as a tree
	\begin{itemize}
		\item Node: Intermediate execution state
		\item Edge: State transition from executing a thread
	\end{itemize}
	\linegap

	State space size depends on which {\bf preemption points} are used.
\end{frame}

@@

	Kernel (``P3'')
	\begin{itemize}
		%\item UNIX-like educational kernel specification
		\item Students implement scheduler, synchronization, VM %from scratch %% trying not to be too arrogant ._.
		\item UNIX-like process lifecycle ({\tt fork}, {\tt exec}, {\tt exit}, {\tt wait})
		%\item Teams of 2
		\item Testing with MC requires manual annotation {\em [Blum '12]}
	\end{itemize}
	\pause
	\linegap

@@

	Results (as of spring 2016)
	\begin{itemize}
		\item 47/90 groups volunteered
		\item 85 concurrency bugs found; 53 which students then fixed
		\item 32/47 found $\ge 1$ concurrency bug, 22/47 fixed $\ge 1$
		\item No visible impact on ultimate grades...
	\end{itemize}
\end{frame}

@@

	Kernel (``threads'')
	\begin{itemize}
		\item Basecode provides round-robin scheduler, synchronization
		\item Students implement priority scheduling, sleep
	\end{itemize}
	\linegap

@@

\begin{frame}{Pintos \& Landslide}
	Existing support:
	\begin{itemize}
		\item Data-race analysis
			%Lockset/vector-clock analyses
			extended to interrupt-disabling
		\item Automatically inserts (most) kernel annotations
		%\item (various engineering challenges) % palloc heap, non-direct-mapped kernelspace, skipping driver init loops
		\item Often requires manual intervention
	\end{itemize}
	% "Works enough for OOPSLA but not enough for the studence"
	\linegap

@@

\begin{frame}{Deliverables}
	In the thesis, I will:
	\begin{itemize}
		\item Pebbles
		\begin{itemize}
			\item Discuss factors keeping Landslide from improving grades
			%\item Discuss how to make P3 and Landslide compromise to work together
			\item Study accuracy of automation heuristics
		\end{itemize}
		\item Pintos
		\begin{itemize}
			\item Open-source port of Landslide to Bochs
			\item 2 semesters of user study in Pintos class(es)
			\item Compare incidence/nature of bugs found to Pebbles
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	MCing a transactional program
	\begin{itemize}
		\item Simulate TM, interleaving transactions, rolling-back upon conflicts
			\begin{itemize}
				\item slow; invites state space explosion
			\end{itemize}
		\item Enforce atomicity via scheduler, anticipate aborts on every transaction
			%via analysis
			%\begin{itemize}
			%	\item (under HTM, all transactions might abort)
			%\end{itemize}
		\item Suffices to enumerate all {\em observationally-equivalent} schedules.
	\end{itemize}
\end{frame}

\begin{frame}{Concurrency model}
	What are a transactional program's {\em observable} behaviours?
	\begin{itemize}
		\item Transactional writes, if successful, appear atomically %to other threads
		\begin{itemize}
			\item Equivalent to a lock protecting that state
		\end{itemize}
			\pause
		\item Simultaneous transactions succeed only if no conflicts
		\begin{itemize}
			\item Equivalent
				%(modulo performance)
				to serialize independent transactions
		\end{itemize}
			\pause
		\item Failing transactions' effects rolled-back with no trace
		\begin{itemize}
			\item Treat each \xbegin as a failure injection point
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Failure Injection - 2 dimensions of concurrency}
	\begin{center}
		\includegraphics[width=\textwidth]{failure-injection-0.pdf}
	\end{center}
\end{frame}
\begin{frame}{Failure Injection - 2 dimensions of concurrency}
	\begin{center}
		\includegraphics[width=\textwidth]{failure-injection-2.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Timeline}
	\begin{itemize}
		\item 2017 Apr-Aug: Port Landslide to Bochs for Pintos
		\item 2017 Apr-Aug: Transactional memory
		\item 2017 Aug-Oct: Pintos user study
		\item 2017 Oct-Nov: Pebbles user study (again)
		\item 2017 Dec:     Start writing thesis
		\item 2018 Jan-Apr: More writing \& miscellany
		\item 2018 Feb-Mar: Pintos \& Pebbles (again)
		\item 2018 Apr-May: Defend \& graduate
	\end{itemize}
\end{frame}

@@

%\begin{frame}{Interleaving t'xns non-atomically $\rightarrow$ state space explosion}
%	% imagine a pair of transactions each with multiple conflicting accesses
%	% they should be treated as atomic
%\end{frame}

%\begin{frame}{STM: Judging when aborts are possible}
%\end{frame}

\begin{frame}{HTM: Sources of Benchmarks}
	\begin{itemize}
		\item Shamay, https://github.com/cmpxchg16/tsx.me, 2014
		\item Dehesa-Azuara and Stanley, http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016
		\item Chapman et al., Hybrid STM/HTM for nested transactions on OpenJDK, OOPSLA 2016
		\item Wang et al., Eunomia: Scaling Concurrent Search Trees under Contention Using HTM, PPoPP 2017
	\end{itemize}
\end{frame}

@@

\section{Warp Zone}

@@

\begin{frame}{Don't you need \texttt{if (mutex\_held(\&foo\_lock)) \_xabort();} in that transaction on slide 31?}
	Yes.
\end{frame}
